How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-17 tse-paper/content.tex
adapt authors
caros for now last text edits

\subsection{Discussion}
Our literature survey sheds light on how build log analysis is used within
computer science research.
Hereafter we will summarize our findings and relate them to the
approach we are taking to improve build log analysis with chunk
retrieval techniques.

We saw that various researchers process build logs to support their
research, often because the logs provide more or the necessary information
to the researchers~\cite{ren2018automated}.
Most of the articles which explicitly mention a manual analysis of
the logs only restrict their work to few logs or take a
``representative sample to make manual analysis
feasible''~\cite{zolfagharinia2017not}.
Automated approaches are necessary as many studies target a large
number of builds.
The chunk retrieval techniques we investigate within this article are
automated, apart from their configuration through training examples.

We have observed very little reuse of build log analysis tools.
This can stem from the fact that publishing replication packages and
research prototypes became popular only in recent years or from the
high specialization of the developed tools in regards to supported
build environments and targeted information.
Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that few build environments are
targeted by many studies and tools (A).
The majority of build environments was however only targeted by one
article each (B).
From this we follow that, while custom parsers are beneficial for the
small group of environments targeted by many (A), all and especially
the rarer investigated environments (B) would benefit from a generic
solution.
We implement build environment agnostic build log analysis
techniques, which are customized to a specific project by
the training examples which the user provides.
In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
data set we chose for our evaluation covers a broad range of build
environments.
% Further, the training examples also specify the log chunk targeted
by the
% retrieval, which enable our chunk retrieval techniques to retrieve a
% broad range of information from build logs.

The articles within our study mainly tried to retrieve or classify
errors or reasons for the build to fail from the logs.
Therefore we chose the log chunk ``describing why the build failed''
~\cite{brandt2020logchunks} for our empirical comparison study
of chunk retrieval techniques.
The techniques we implement are in principle agnostic to the targeted
log chunk, thus we believe our results can be generalized to other
kinds of log chunks.

Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone'', other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.
For this reason, we are focussing on techniques that are configured
through examples consisting of logs and corresponding chunks.
These examples are provided by the user and can simply be
exchanged whenever new cases are introduced.

@@

In summary, this literature study shows the need for build log
analysis techniques that are build-environment agnostic, automated
and can be configured by the user with little effort.
We achieve this by developing three techniques for which the
information to be retrieved is specified through examples or keywords.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log.
RLR mimics the
situation of guessing blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.

@@

\subsection{Chunk Retrieval Example}
\begin{figure*}[!htbp]
  \centering
\begin{subfigure}[!htbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

@@

To illustrate the chunk retrieval techniques further, we want to give
an example of real-world log chunks and how the techniques can
retrieve them from build logs.
\Cref{lst:chunk-example} shows two exemplary
log chunks from Android build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed, the information most often targeted by
other articles which analyze build logs.

When PBE is trained with these two examples it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators, consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if the third example from \Cref{lst:chunk-example-3} with a
different structural representation is added to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}''.
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of our empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

CTS will rank the lines from the analyzed build log which are most
similar to the chunks in the training examples,
marked blue in \Cref{lst:chunk-example}.
It will then return the two lines ranked highest, as the chunks in
the training examples have on average two lines.

KWS is configured with keywords that appear close to the chunk within
the build log.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE'', ``wrong'' or ``failed''.
The example from \Cref{lst:chunk-example-3} could be found with
``FAIL''.
If KWS would be trained with all three examples and the afformentioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice, most often within
the training set of three examples.

\subsection{Tool Implementation}
For our comparison study we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
The implementation of PBE in C\# is based on the Microsoft PROSE
library~\cite{prose2019webpage}.
We implemented CTS, KWS and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.

@@

To investigate when our chunk retrieval techniques are suited to
analyze build logs (\textbf{RQ2}), we perform an empirical comparison
study on the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
In particular we are interested in how many examples a technique
should be trained with for best performance (\textbf{RQ2.1}),
how similar the structural representation of the targeted log chunks
has to be (\textbf{RQ2.2}) and how reliable the quality of the
produced outputs are (\textbf{RQ2.3}).

@@

\subsection{LogChunks}
For this study we created the previously published data set
\emph{LogChunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, the authors manually ``labeled
the log part (chunk) describing why the build
failed''~\cite{brandt2020logchunks}.
We included keywords, which we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log we assigned them the same structural category.

\subsection{Study Process}
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---end therefore the
same build environment (1).
One to five examples make up the training set (2).
We vary its size to measure how many examples are needed to
confidently configure a technique.
The training set is so small, because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort.
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a training set size of 1, as users of our techniques
should receive reliable results on every build log they analyze.
As a next step, we configure the chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
  \item The size of training set (\textbf{RQ2.1}),
  \item the number of structural categories in the training set
  (\textbf{RQ2.2})
  \item the lines of the retrieval output ($\mathit{RetrievedLines}$),
  and our oracle, the output defined in the test example
  ($\mathit{DesiredLines}$).
  From these we calculate accuracy metrics (\textbf{RQ2.1}).
\end{itemize}

\textbf{Accuracy Metrics}
We calculate a number of accuracy metrics to measure how well
the chunk retrieval techniques perform in our evaluation:

>>>>

\subsection{Discussion}
\label{sec:lit-sur:discussion}
Our literature survey sheds light on how build log analysis is used within
computer science research.
Hereafter we will summarize our findings and relate them to the
approach we are taking to improve build log analysis with chunk
retrieval techniques.

We saw that various researchers process build logs to support their
research, often because the logs provide more or the necessary information
to the researchers~\cite{ren2018automated}.
Articles which explicitly mention a manual analysis of
the logs, mostly restrict their work to few logs or take a
``representative sample to make manual analysis
feasible''~\cite{zolfagharinia2017not}.
Automated approaches are necessary as many studies target a large
number of builds.
% The chunk retrieval techniques we investigate within this article are
% automated, apart from their configuration through training examples.

We have observed very little reuse of build log analysis tools.
This can stem from the fact that publishing replication packages and
research prototypes became popular only in recent years or from the
high specialization of the developed tools in regards to supported
build environments and targeted information.

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that few build environments are
targeted by many studies and tools (A).
The majority of build environments was however only targeted by one
article each (B).
From this we follow that, while custom parsers are beneficial for the
small group of environments targeted by many (A), all and especially
the rarer investigated environments (B) would benefit from a generic
solution.
% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone'', other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.

@@

In summary, this literature study shows the need for build log
analysis techniques that are automated, build-environment agnostic
and can be configured by the user with little effort.
% We achieve this by developing three techniques for which the
% information to be retrieved is specified through examples or keywords.

For this reason, we are focussing on techniques that run automatic
after being configured through examples or keywords.
Examples consist of logs and corresponding chunks.
The examples and keywords are provided by the user and can simply be
exchanged whenever new cases are introduced.

@@

% The articles within our study mainly tried to retrieve or classify
% errors or reasons for the build to fail from the logs.
% Therefore we chose the log chunk ``describing why the build failed''
% ~\cite{brandt2020logchunks} for our empirical comparison study
% of chunk retrieval techniques.
% The techniques we implement are in principle agnostic to the targeted
% log chunk, thus we believe our results can be generalized to other
% kinds of log chunks.

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
In addition we present our comparison baseline of random line
retrieval (RLR), illustrate the techniques on a concrete example
of log chunks and sketch out implementation.

@@

\subsection{Random Line Retrieval (RLR)}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log.
RLR mimics the
situation of guessing blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.

@@

\subsection{Chunk Retrieval Example}
\label{seC:crt-example}

\begin{figure*}[!htbp]
  \centering
\begin{subfigure}[!htbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

@@

To illustrate the chunk retrieval techniques further, we want to give
an example of real-world log LogChunks.
\Cref{lst:chunk-example} shows two excerpts
from Android build logs.
Marked in blue are the chunks
which contain the information on why the
corresponding build failed.

When PBE is trained with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if the third example from \Cref{lst:chunk-example-3},
which has a
different structural representation, is added to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}''.
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of our empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

CTS will rank the lines from an analyzed build log which are most
similar to the chunks in the training examples
from \Cref{lst:chunk-example}.
It will then return the two lines ranked highest, as the chunks in
the training examples have on average two lines.

KWS is configured with keywords that appear close to the chunk within
the build log.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE'', ``wrong'' or ``failed''.
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL''.
If KWS would be trained with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

\subsection{Tool Implementation}
For our comparison study we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
The implementation of PBE in C\# is based on the Microsoft PROSE
library~\cite{prose2019webpage}.
We implemented CTS, KWS and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.
They can be found in our replication
package~\cite{brandt2020chunk-replication} together which have
our techniques installed and ready to use.

@@

To investigate when our chunk retrieval techniques are suited to
analyze build logs (\textbf{RQ2}), we perform an empirical comparison
study on the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
In particular we are interested in how many examples a technique
should be trained with (\textbf{RQ2.1}),
how similar the structural representation of the targeted log chunks
has to be (\textbf{RQ2.2}) and how reliable the quality of the
produced outputs is (\textbf{RQ2.3}).

@@

\subsection{LogChunks}
For this study we created the previously published data set
\emph{LogChunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, the authors manually ``labeled
the log part (chunk) describing why the build
failed''~\cite{brandt2020logchunks}.
We included keywords, which we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

\subsection{Study Process}
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---and thus the
same build environment (1).
One to five examples make up the training set (2).
We vary its size to measure how many examples are needed to
confidently configure a technique.
The training set is so small, because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a test set size of 1, as users of our techniques
should receive reliable results on every build log they analyze.
In the next step, we configure a chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
the size of training set (\textbf{RQ2.1}),
the number of structural categories in the training
set (\textbf{RQ2.2}),
the lines of the retrieval output ($\mathit{RetrievedLines}$),
and our oracle: the output defined in the test example
($\mathit{DesiredLines}$).
From these we calculate accuracy metrics (\textbf{RQ2.3}):
